Epistemic Semantics in Guarded String Models

### 2020-06-18 tex/draft.tex
update guarded string length

Guarded strings are strings of events, alternating with such atoms,
and starting and ending with atoms.  \ex{gs1} gives the encoding as
guarded strings of the worlds in \ex{elevator1} and \ex{coin1}.

%% \begin{example}   \label{gs1}
%% \begin{tabular}[t]{ll}
%% $v_1$ & $\bar p \bar q ~u~ \bar p q ~u~ q \bar p ~d~ \bar p q ~u~ q \bar p$ \\
%% $v_2$ & $\bar p \bar q ~u~ \bar p q ~u~ p\bar q ~u~ pq$ \\   
%% $w_1$ & $\bar h t$ \\
%% $w_2$ & $\bar h t ~a_0~ \bar h t$ \\  
%% $w_3$ & $\bar h t ~a_0~ \bar h t ~b_0~ \bar h t$ \\
%% \end{tabular}
%% \end{example}
\begin{example}   \label{gs1}
\begin{tabular}[t]{ll}
$v_1$ & $\hat 0 ~u~ \hat 1 ~u~ \hat 2 ~d~ \hat 1 ~u~ \hat 2$ \\
$v_2$ & $\hat 0 ~u~ \hat 1 ~u~ \hat 2 ~u~ \hat 3$ \\   
$w_1$ & $H$ \\
$w_2$ & $H ~a_0~ H$ \\  
$w_3$ & $T~a_0~ T ~b_0~ T$ \\
\end{tabular}
\end{example}

@@

We also implement the semantics of Epik terms using lazy lists~\cite{LazyLists}
in Haskell, rather than the direct interpretation as sets. Unfortunately,
regular expressions, and hence also Epik, can denote infinite sets of strings,
for example the term $\_^*$. Normally regular languages are represented using a
finite coalgebra~\cite{}. However the non-distributivity of $\diamond$ accross
$;$ complicates this construction~\footnote{Axiomatic and Coalgebraic models for
  Epik are open questions}. To sidestep this, we parameterize the interpretation
function on a positive integer $n$ and only produce guarded strings of length
$n$ or less.  The length of a guarded string $g$, written $|g|$ is the number of
events in the string, so $\alpha$ has length $0$, $\alpha e\beta$ has length
$1$, and $\alpha e \beta e' \gamma$ has length $2$.

We translate the Epik terms into a Haskell algebraic datatype that represents
terms with the same signature as described in Section 2. We then paramaterize
the interpretation function, $\llparenthesis p \rrparenthesis^{\varphi,{\mathcal{E}}}_n$ on an
integer $n$ that describes the maximum length of a string we will produce. The
bounded interpretation into lists of strings is very similar to the unbounded
interpretation into sets of strings, except for the bounds checking. The full
details are shown in Figure~\ref{fig:list-interp}: we use list-comprehension
notation, which is analogous to set-builder notation, except that it produces a
list (potentially with duplicated elements) instead of a set, and so is written
with square brackets $[]$. We use $+$ to indicate list concatenation, and $\sum
l$ to indicate the concatenation of a list of lists $l$. We overload the set
membership symbol $\in$ to also indicate membership in a list. We also lift the
fusion operator $\cdot$ to lists. For a list of guarded strings $l$, we write
$l|_n$ as shorthand for the set $[ g~|~g \in l , |g| \leq n]$.

@@

The final step is to convert the list of guarded strings into a set of guarded
strings. Since a list is just an ordered multiset, we forget the order, and
remove all duplicates. Haskell's libraries have a built-in function for this,
\texttt{Data.Set.fromList}, which converts a list into a set.

\begin{figure}
  \[\begin{array}{r >{\triangleq}c l}
  \llparenthesis p \rrparenthesis^{\varphi,{\mathcal{E}}}_0 && [] \\
  \llparenthesis 0 \rrparenthesis^{\varphi,{\mathcal{E}}}_n && [] \\
  \llparenthesis 1 \rrparenthesis^{\varphi,{\mathcal{E}}}_n && [\alpha \mid \alpha \leq \varphi] \\
  \llparenthesis e \rrparenthesis^{\varphi,{\mathcal{E}}}_n && [\alpha \cdot e \cdot \beta \mid (\alpha, \beta) \in {\mathcal{E}}(e), \alpha\vee \beta \Rightarrow \varphi] \\
  \llparenthesis \_ \rrparenthesis^{\varphi,{\mathcal{E}}}_n && \sum [\llparenthesis e \rrparenthesis^{\varphi,{\mathcal{E}}} \mid e \in \dom({\mathcal{E}}) ] \\
  \llparenthesis t \rrparenthesis^{\varphi,{\mathcal{E}}}_n && [\alpha\mid \alpha \Rightarrow \varphi \wedge t] \\
  \llparenthesis p + q \rrparenthesis^{\varphi,{\mathcal{E}}}_n && \llparenthesis p \rrparenthesis^{\varphi,{\mathcal{E}}}_n + \llparenthesis q \rrparenthesis^{\varphi,{\mathcal{E}}}_n \\
  \llparenthesis p \& q \rrparenthesis^{\varphi,{\mathcal{E}}}_n &&
  [g~|~g \in \llparenthesis p \rrparenthesis_n^{\varphi,{\mathcal{E}}}, g \in \llparenthesis q \rrparenthesis_n^{\varphi,{\mathcal{E}}}] \\
  \llparenthesis p;q \rrparenthesis^{\varphi,{\mathcal{E}}}_n &&
  \left(\llparenthesis p \rrparenthesis^{\varphi,{\mathcal{E}}}_n \cdot \llparenthesis q\rrparenthesis^{\varphi,{\mathcal{E}}}_n\right)|_n\\
  \llparenthesis p^* \rrparenthesis^{\varphi,{\mathcal{E}}}_n && [] + \left(\llparenthesis p\rrparenthesis^{\varphi,{\mathcal{E}}}_n \cdot  \llparenthesis p^* \rrparenthesis^{\varphi,{\mathcal{E}}}_{n-i}\right)|_n \\
  \multicolumn 2 r {} & \text{where}~i = \max\{1,\min\{|g|\mid g \in \llbracket p \rrbracket^{\varphi,{\mathcal{E}}}\}\}\\
  \llparenthesis \sim p \rrparenthesis^{\varphi,{\mathcal{E}}}_n && [g \mid g \in \llbracket \_^* \rrbracket^{\varphi,{\mathcal{E}}}_n, g \not\in \llbracket p \rrbracket^{\varphi,{\mathcal{E}}}_n] \\
  \llparenthesis \langle a \rangle p \rrparenthesis^{\varphi,{\mathcal{E}}}_n &&
  [g' \mid g \in g'~\doublehat{R_a}~g, g \in \llparenthesis p \rrparenthesis^{\varphi,{\mathcal{E}}}]
  \end{array}
  \]
\end{figure}

>>>>

Guarded strings are strings of events, alternating with such atoms,
and starting and ending with atoms. The length of a guarded string
$g$, written $|g|$ is the number of events in $g$, ignoring the atoms.
\ex{gs1} gives the encoding as guarded strings of the worlds in
\ex{elevator1} and \ex{coin1}.

%% \begin{example}   \label{gs1}
%% \begin{tabular}[t]{ll}
%% $v_1$ & $\bar p \bar q ~u~ \bar p q ~u~ q \bar p ~d~ \bar p q ~u~ q \bar p$ \\
%% $v_2$ & $\bar p \bar q ~u~ \bar p q ~u~ p\bar q ~u~ pq$ \\   
%% $w_1$ & $\bar h t$ \\
%% $w_2$ & $\bar h t ~a_0~ \bar h t$ \\  
%% $w_3$ & $\bar h t ~a_0~ \bar h t ~b_0~ \bar h t$ \\
%% \end{tabular}
%% \end{example}
\begin{example}   \label{gs1}
  \renewcommand{\arraystretch}{1.2}
  \begin{tabular}[t]{llc}
     \multicolumn 2 c {\textit{Guarded String}} & \textit{Length}\\ \hline
$v_1$ & $\hat 0 ~u~ \hat 1 ~u~ \hat 2 ~d~ \hat 1 ~u~ \hat 2$ & $4$ \\
$v_2$ & $\hat 0 ~u~ \hat 1 ~u~ \hat 2 ~u~ \hat 3$ & $3$ \\   
$w_1$ & $H$ & $0$\\
$w_2$ & $H ~a_0~ H$ & $1$ \\  
$w_3$ & $T~a_0~ T ~b_0~ T$ & $3$\\
  \end{tabular}
    \renewcommand{\arraystretch}{1}
\end{example}

@@

We also implement the semantics of Epik terms using lazy
lists~\cite{LazyLists} in Haskell, rather than the direct
interpretation as sets. Unfortunately, regular expressions, and hence
also Epik, can denote infinite sets of strings, for example the term
$\_^*$. Normally regular languages are represented using a finite
coalgebra~\cite{}. However the non-distributivity of $\diamond$
accross $;$ complicates this construction\footnote{Axiomatic and
  Coalgebraic models for Epik are open questions}. To sidestep this,
we parameterize the interpretation function on a positive integer $n$
and only produce guarded strings of length $n$ or less.

We translate the Epik terms into a Haskell algebraic datatype that
represents terms with the same signature as described in Section 2. We
then paramaterize the interpretation function, $\llparenthesis p
\rrparenthesis^{\varphi,{\mathcal{E}}}_n$ on an integer $n$ that
describes the maximum length of a string we will produce. The bounded
interpretation into lists of strings is very similar to the unbounded
interpretation into sets of strings, except for the bounds
checking. The full details are shown in Figure~\ref{fig:list-interp}:
we use list-comprehension notation, which is analogous to set-builder
notation, except that it produces a list (potentially with duplicated
elements) instead of a set, and so is written with square brackets
$[]$. We use $+$ to indicate list concatenation, and $\sum l$ to
indicate the concatenation of a list of lists $l$. We overload the set
membership symbol $\in$ to also indicate membership in a list. We also
lift the fusion operator $\diamond$ to lists. For a list of guarded
strings $l$, we write $l|_n$ as shorthand for the set $[ g~|~g \in l ,
  |g| \leq n]$. Finally, we write $\lfloor X \rfloor$ to convert a set
$X$ into a list containing same elements with an arbitrary order.

@@

To convert from a list of guarded strings $l$ to a set of guarded
strings, we simply write $\lceil l\rceil$. Note that for any $p$,
$\varphi$, $\mathcal E$, and $n$, $\left\lceil\llparenthesis p
\rrparenthesis^{\varphi,\mathcal E}_n\right\rceil = \{g \mid g \in
\llbracket p \rrbracket^{\varphi,\mathcal E}, |g| \leq n\}$.

\begin{figure}
  \[\begin{array}{r >{\triangleq}c l}
  \llparenthesis p \rrparenthesis^{\varphi,{\mathcal{E}}}_0 && [] \\
  \llparenthesis 0 \rrparenthesis^{\varphi,{\mathcal{E}}}_n && [] \\
  \llparenthesis 1 \rrparenthesis^{\varphi,{\mathcal{E}}}_n && \lfloor \mathcal A_{\mathsf B}^\varphi \rfloor \\
  \llparenthesis e \rrparenthesis^{\varphi,{\mathcal{E}}}_n && \lfloor \hat{\mathcal E}^\varphi(e)\rfloor \\
  \llparenthesis \psi \rrparenthesis^{\varphi,{\mathcal{E}}}_n && \lfloor \mathcal A_{\mathsf B}^{\varphi~\psi} \rfloor \\
  \llparenthesis p + q \rrparenthesis^{\varphi,{\mathcal{E}}}_n && \llparenthesis p \rrparenthesis^{\varphi,{\mathcal{E}}}_n + \llparenthesis q \rrparenthesis^{\varphi,{\mathcal{E}}}_n \\
  \llparenthesis p;q \rrparenthesis^{\varphi,{\mathcal{E}}}_n &&
  \left(\llparenthesis p \rrparenthesis^{\varphi,{\mathcal{E}}}_n \cdot \llparenthesis q\rrparenthesis^{\varphi,{\mathcal{E}}}_n\right)|_n\\
  \llparenthesis p^* \rrparenthesis^{\varphi,{\mathcal{E}}}_n && [] + \left(\llparenthesis p\rrparenthesis^{\varphi,{\mathcal{E}}}_n \diamond  \llparenthesis p^* \rrparenthesis^{\varphi,{\mathcal{E}}}_{n-i}\right)|_n \\
  \multicolumn 2 r {} & \text{where}~i = \max\{1,\min\{|g|\mid g \in \llbracket p \rrbracket^{\varphi,{\mathcal{E}}}\}\}\\
  \llparenthesis \neg p \rrparenthesis^{\varphi,{\mathcal{E}}}_n && [g \mid g \in \llbracket \_^* \rrbracket^{\varphi,{\mathcal{E}}}_n, g \not\in \llbracket p \rrbracket^{\varphi,{\mathcal{E}}}_n] \\
  \llparenthesis \Diamond_a~p \rrparenthesis^{\varphi,{\mathcal{E}}}_n &&
  [g' \mid g \in g'~\hat{R_a}~g, g \in \llparenthesis p \rrparenthesis^{\varphi,{\mathcal{E}}}]
  \end{array}
  \]
  \caption{Bounded interpretation using lazy lists}
\end{figure}
